---
title: Lobster Abundance and Size Analysis"
author: "Emma"
format: html
---

## Exercise: Explore, Clean, and Wrangle Data

```{r}
#| warning: false
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(here)
library(DT)
```

```{r}
#| warning: false
#Read in data
lobster_abundance <- read_csv("../data/Lobster_Abundance_All_Years_20220829.csv")
datatable(lobster_abundance)
```

```{r}
#Explore Data, High Level Summary
head(lobster_abundance)

summary(lobster_abundance)

str(lobster_abundance)

colnames(lobster_abundance)

```
1. The variable SIZE_MM uses -99999 as the code for missing values (see metadata or use unique()). This has the potential to cause conflicts with our analyses, so let’s convert -99999 to an NA value. Do this using mutate() and na_if(). Look up the help page to see how to use na_if(). Check your output data using unique().
```{r}
# Fix NA Vals
lobster_abundance <- lobster_abundance |> 
  mutate(SIZE_MM == na_if(SIZE_MM,-99999))
 head(lobster_abundance)
```

2. Create and store a subset that does NOT include observations from Naples Reef (NAPL). Check your output data frame to ensure that NAPL is NOT in the data frame.
```{r}
#Lobsters Not in NAPL
not_napl <- lobster_abundance |> 
  filter(SITE != "NAPL")

head(not_napl)
```

3. Create and store a subset with lobsters at Arroyo Quemado (AQUE) AND with a carapace length greater than 70 mm. Check your output
```{r}
aque_70mm <- lobster_abundance |> 
  filter(SITE == "AQUE" & SIZE_MM > 70)
head(aque_70mm)
```
4. Find the maximum carapace length using max() and group by SITE and MONTH. Think about how you want to treat the NA values in SIZE_MM (Hint: check the arguments in max()). Check your output.

```{r}
max_lobster <-  lobster_abundance |> 
  group_by(SITE, MONTH) |> 
  summarise(MAX_LENGTH = max(SIZE_MM, na.rm = TRUE))
max_lobster 
```

## Exercise: Data Visualization

```{r}
lobster_abundance_cleaned <- lobster_abundance |> 
  filter(SIZE_MM != -99999 & !is.na(SIZE_MM))
lobster_abundance_cleaned 

```

5. Create a multi-panel plot of lobster carapace length (SIZE_MM) using ggplot(), geom_histogram(), and facet_wrap(). Use the variable SITE in facet_wrap(). Use the object lobster_abundance.

```{r}
ggplot(data = lobster_abundance_cleaned, 
       aes(x = SIZE_MM)) +
    geom_histogram() +
    facet_wrap(~SITE)
```

6. Create a line graph of the number of total lobsters observed (y-axis) by year (x-axis) in the study, grouped by SITE.

```{r}
lobsters_summarize <- lobster_abundance_cleaned |> 
  group_by(SITE, YEAR) |> 
  summarize(COUNT = n())

lobsters_summarize
```

```{r}
ggplot(lobsters_summarize,
       aes(x = YEAR, y= COUNT, color = SITE)) +
  geom_line() +
  geom_point()
```

7. Create a bar graph that shows the amount of small and large sized carapace lobsters at each SITE from 2019-2021. Note: The small and large metrics are completely made up and are not based on any known facts.


```{r}
lobster_size_lrg <- lobster_abundance_cleaned |> 
  filter(YEAR %in% c(2019, 2020, 2021)) |> 
  mutate(SIZE_BIN = if_else(SIZE_MM <= 70, "small", "large")) |> 
  group_by(SITE, SIZE_BIN) |> 
  summarize(COUNT = n())

lobster_size_lrg

```
```{r}
# bar plot
ggplot(data = lobster_size_lrg, aes(x = SITE, y = COUNT, fill = SIZE_BIN)) +
    geom_col()

# dodged bar plot
ggplot(data = lobster_size_lrg, aes(x = SITE, y = COUNT, fill = SIZE_BIN)) +
    geom_col(position = "dodge")

```




