---
title: Lobster Abundance and Size Analysis"
author: "Emma"
format: html
---

## Exercise: Explore, Clean, and Wrangle Data

```{r}
#| warning: false
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(here)
library(DT)
```

```{r}
#| warning: false
#Read in data
lobster_abundance <- read_csv("../data/Lobster_Abundance_All_Years_20220829.csv")
datatable(lobster_abundance)
```

```{r}
#Explore Data, High Level Summary
head(lobster_abundance)

summary(lobster_abundance)

str(lobster_abundance)

colnames(lobster_abundance)

```
1. The variable SIZE_MM uses -99999 as the code for missing values (see metadata or use unique()). This has the potential to cause conflicts with our analyses, so let’s convert -99999 to an NA value. Do this using mutate() and na_if(). Look up the help page to see how to use na_if(). Check your output data using unique().
```{r}
# Fix NA Vals
lobster_abundance <- lobster_abundance |> 
  mutate(SIZE_MM == na_if(SIZE_MM, -99999))
```

2. Create and store a subset that does NOT include observations from Naples Reef (NAPL). Check your output data frame to ensure that NAPL is NOT in the data frame.
```{r}
#Lobsters Not in NAPL
not_napl <- lobster_abundance |> 
  filter(SITE != "NAPL")

head(not_napl)
```

3. Create and store a subset with lobsters at Arroyo Quemado (AQUE) AND with a carapace length greater than 70 mm. Check your output
```{r}
aque_70mm <- lobster_abundance |> 
  filter(SITE == "AQUE" & SIZE_MM > 70)
head(aque_70mm)
```
4. Find the maximum carapace length using max() and group by SITE and MONTH. Think about how you want to treat the NA values in SIZE_MM (Hint: check the arguments in max()). Check your output.

```{r}
max_lobster <-  lobster_abundance |> 
  group_by(SITE, MONTH) |> 
  summarise(MAX_LENGTH = max(SIZE_MM, na.rm = TRUE))


max_lobster 
```

